Learning to Synthesize Rules for Network Intrusion Detectors

### 2020-01-15 paper/paper.tex
pass

\subsection{\Fix{JBoss JMX Console Beanshell Deployer WAR Upload and Deployment Exploit}}
\label{sec:content-example}

@@

\section{Technique}

@@

@@

\subsubsection{Frequency of content modifiers.} As previously
mentioned, \percRulesWithContent\ of Suricata rules involve content
options. Furthermnore, we observed that many of these rules indicate
what parts of the message the content string should be checked
against.\Mar{Lucas, (1) vc. tem dados para dar suporte a isto (de http
  e tcp)? ou seja, quantas options do tipo ``content'' possuem
  modificadores.}  For example, the creator of the rule may want to
indicate that a specific string be present in the header of an HTTP
request message. Content \emph{modifiers} serve this purpose. For
example, the modifier \CodeIn{http\_header} in the option
\CodeIn{content: ``keep-alive''; http\_header} states the requirement that
the string ``keep-alive'' be present in the header of the message. (A
keep-alive connection attribute indicates that a single TCP connection
should remain open for multiple HTTP requests/responses.)
Figure~\Fix{adicione pequena figura...verifique se ainda faz sentido
  depois do que falei acima...}

Suricata supports content modifiers for tcp and http. Given that
content options often use modifiers\Mar{assumindo que eh
  verdade. confirmar estat. solicitada acima...}, \tname\ uses
associated frequency to discriminate more or less likely rules. For
example, historical data indicates that content options that use the
\CodeIn{http\_connection} modifiers are much less likely than content
options that use the modifier
\CodeIn{http\_uri}. Figure~\ref{fig:distribution-content_modifiers}
shows the distribution of content modifiers for http. We ommit the
distribution for tcp for space. Similarly to the the previous
heuristic functions, this function uses the frequency of content
modifiers to calibrate the contribution of each rule.

@@

@@

\subsection{Pontuação}
A avaliação que fazemos aqui pontua um dado content de acordo com quantas vezes o campo em que ele se encontra eh referenciado nas regras. Por exemplo, vamos supor que o tokenizer produziu a opção content: "/HtmlAdaptor". Essa string pertence ao campo "Request URI:". Esse campo pode ser acessado pelo modificador http\_uri. O tokenizer associa o content a esse modificador. Posteriormente, esse content vai receber uma pontuação correspondente a popularidade desse modificador:\\
\[\frac{10111}{10111*1}\]
A multiplicação por 1 no denominador eh referente a quantidade de modificadores que estão associadas a esse content.
No caso da opção content: "keep-alive", essa string pertence ao campo "Connection:". Como falado acima, o suricata pode acessar esse campo pelos modificadores http\_header ou http\_connection. Portanto, esse content será pontuado de acordo com a popularidade desses dois modificadores:
\[\frac{4089+29}{10111*2}\]

Content options are
associated with specific packets fields\Mar{or modifiers?}.\Mar{why
  GET is associated with http\_method and http\_header?}\Mar{are all
  content associated with one or many fields or positions in a packet?
  can you give examples other than GET to explain this? if yes,
  shouldn't this be discussed before? Maybe Section 2.3?}\Mar{Also, it
  is important to justify (to explain rationale)---how is it different
  from counting rarity?}

@@

More precisely, we calculate the contribution of the frequency of
content modifiers with the following formula:

\[\frac{\sum_{i=1}^{N}\sum_{j=1}^{M}\frac{modifier\_frequency[modifier_{ij}]}{max(modifier\_frequency)}}{N}\]

\noindent
, where $N$ is the number of contents in a given rule and $M$
is the number of content modifiers\Mar{in a rule? what is a content
  modifier in a rule?}. \Luc{M eh o numero de modificadores associados
  a cada content. Temos content[1], content[2], ..., content[N]. Cada
  um desses contents tem modificadores associados content[i][1],
  content[i][2], ..., content[i][M].}

>>>>

\subsection{\Fix{JBoss JMX Console Beanshell Deployer WAR Upload and Deployment Exploit}}
\label{sec:jboss}

@@

\section{Technique}
\label{sec:technique}

@@

\Mar{Revisar.........................}
\Fix{
A avaliação que fazemos aqui pontua um dado content de acordo com
quantas vezes o campo em que ele se encontra eh referenciado nas
regras. Por exemplo, vamos supor que o tokenizer produziu a opção
content: "/HtmlAdaptor". Essa string pertence ao campo "Request
URI:". Esse campo pode ser acessado pelo modificador http\_uri. O
tokenizer associa o content a esse modificador.
}

@@

\subsubsection{Frequency of content modifiers.} As previously
mentioned, \percRulesWithContent\ of Suricata rules involve content
options. Furthermore, we observed---by analyzing the dataset of
existing rules (see Section~\ref{sec:dataset-benign})---that often
these rules indicate what part of the message the content string
should be checked against.\Mar{Lucas, (1) vc. tem dados para dar
  suporte a isto (de http e tcp)? ou seja, quantas options do tipo
  ``content'' possuem modificadores.} \Luc{conseguirei esses dados}
For example, the creator of the rule may want to indicate that a
specific string be present in the header of an HTTP request
message. Content \emph{modifiers} serve this purpose. For example, the
modifier \CodeIn{http\_header} in the option \CodeIn{content:
  ``keep-alive''; http\_header} states the requirement that the string
``keep-alive'' be present in the header of the message. (A keep-alive
connection attribute indicates that a single TCP connection should
remain open for multiple HTTP requests/responses.)
Figure~\Fix{adicione pequena figura...verifique se ainda faz sentido
  depois do que falei acima...}

%% Suricata supports content modifiers for tcp and http. \Luc{Eh possivel utilizar os modificadores em qualquer protocolo. Nós consideramos apenas tcp e http pois são os protocolos que possuem payloads legiveis, assim conseguimos saber melhor como tokenizar o conteudo. O payload de pacotes de outros protocolos, em geral, são um conjunto de bytes sem significado para nós. Existem diversos outros modificadores q não estamos considerando e q tambem permitem procurar numa determinada posição para os casos em q o conteudo do pacote n eh legivel, não citei pq não suportamos isso. Se necessário posso explicar.} Given that
%% content options often use modifiers\Mar{assumindo que eh
%%   verdade. confirmar estat. solicitada acima...},

\tname\ uses the frequency of content modifiers to discriminate more
or less likely rules. For example, historical data indicates that
content options that use the \CodeIn{http\_connection} modifier are
much less likely than content options that use the modifier
\CodeIn{http\_uri}. Figure~\ref{fig:distribution-content_modifiers}
shows the histogram of content modifiers for http. Although content
modifiers are applicable for rules of any protocol in Suricatax,
currently, this heuristic function is only supported for tcp and http
in \tname. The rationale is that (1) these are the protocols involving
most attacks \Mar{Lucas, please confirm.} and (2) the payload of
messages of different protocols are not as segmented as the case of
tcp and http, which makes the reference to specific parts of the
payload less common for rule writers and more challenging for
\tname\Mar{Lucas, please confirm}.

Similarly to the the previous heuristic functions, this function uses
the frequency of content modifiers to calibrate the contribution of
each rule. More precisely, we calculate this heuristic function with
the following formula:

\[\frac{\sum_{i=1}^{N}\sum_{j=1}^{M}\frac{\mathit{modifier\_frequency[modifier_{ij}}]}{\mathit{max(modifier\_frequency)}}}{N}\]

\noindent
, where $N$ is the number of contents in a given rule and $M$ is the
number of content modifiers\Mar{isto eh confuso. vc. esta dizendo que
  o mesmo content pode ter mais de um modificador? ou a string poderia
  ser acessada por mais de um modificador. se for este o caso,
  sinceramente, achei isto bem confuso. afinal, a regra jah indica
  qual o modificador foi escolhido para acessar a string.}. \Luc{M eh
  o numero de modificadores associados a cada content. Temos
  content[1], content[2], ..., content[N]. Cada um desses contents tem
  modificadores associados content[i][1], content[i][2], ...,
  content[i][M].}\Mar{Lucas, e se nao houver modificador algum em um
  content. a frequencia seria 0, correto?}

@@

\Mar{Lucas, use uma regra concreta com a regra da Secao 3.3 (assumindo
que vai mante-la) e mostre como eh feito o calculo da regra desta
funcao heuristica para regra inteira.}

Consider a rule containing the option \CodeIn{content:
  ``/HtmlAdaptor''; http\_uri}, as the one from
Section~\ref{sec:jboss}. The contribution of this option to the score
of the corresponding rule would be $\frac{10,111}{10,111*1}$. The
value of $N$ for this case is 1 as the field of the HTTP message
\CodeIn{Request URI} holding the string can be accessed only through
the modifier \CodeIn{http\_uri}.

%% A avaliação que fazemos aqui pontua um dado content de acordo com quantas vezes o campo em que ele se encontra eh referenciado nas regras. Por exemplo, vamos supor que o tokenizer produziu a opção content: "/HtmlAdaptor". Essa string pertence ao campo "Request URI:". Esse campo pode ser acessado pelo modificador http\_uri. O tokenizer associa o content a esse modificador. 
%% Posteriormente, esse content vai receber uma pontuação
%% correspondente a popularidade desse modificador:\\

@@

A multiplicação por 1 no denominador eh referente a quantidade de
modificadores que estão associadas a esse content.\Mar{nao foi isto
  que vc. disse acima, Lucas. foi dito que N (=2) era o numero de
  contents e o exemplo acima tem apenas 1 content.}  No caso da opção
content: "keep-alive", essa string pertence ao campo
"Connection:". Como falado acima, o suricata pode acessar esse campo
pelos modificadores http\_header ou http\_connection. Portanto, esse
content será pontuado de acordo com a popularidade desses dois
modificadores:
\[\frac{4089+29}{10111*2}\]

%% Content options are
%% associated with specific packets fields\Mar{or modifiers?}.\Mar{why
%%   GET is associated with http\_method and http\_header?}\Mar{are all
%%   content associated with one or many fields or positions in a packet?
%%   can you give examples other than GET to explain this? if yes,
%%   shouldn't this be discussed before? Maybe Section 2.3?}\Mar{Also, it
%%   is important to justify (to explain rationale)---how is it different
%%   from counting rarity?}

@@
